Fast Sorting on Modern Computers

### 2005-08-18 docs/tech_report/conclusions.tex
added quicskort conclusions

TODO quick

TODO radix

>>>>

Sedgewick proposed removing individual insertion sorts in quicksort with one
elegant sweep at the end. LaMarca proposed replacing the large number of
insertion sorts in order to reduce cache misses. We note that attempting to move
the pivots, as Sedgewick does, actually increases both the number of branches
and the rate of mispredictions. We conclude that LaMarca's improvment, which
slightly improves quicksorts cache performance, also slightly improves it's
branch prediction performance.

However, his move to multi-quicksort, while not affecting the rate of branch
mispredictions in the partitioning phase, results in a large increase in branch
mispredictions due to the addition of a binary search. We determine that this
while reducing the instruction count, the increase is actual running time due to
branch mispredictions is significant.

We also analysed the effects of different medians on quicksort. We deterine that
not using a median to choose a pivot makes the branches more predictable than
they are biased, though at a cost of a large increase in instructions (and also
significantly greater possability of a worst-case $O(N^2)$ sort). We also saw
that the more keys considered when determining the median, the lower the number
of branches. However, this can result in an increase in the rate of branch
mispredictions. We determine that reducing entropy is not the best way of making
quicksort perform faster, and that using a median-of-three actually achieves the
higher performance.

As predicted, radixsort is highly predictable, having no comparison
mispredictions at all. This result, combined with its linear complexity, and its
improved cache performance, mean that it is the fastest of any of the sorts
presented here.
